Light inextensible strings under tension in the Schwarzschild geometry

### 2019-08-28 string_physics_teacher.tex
slight rewording

A light string's being under tension means that its path will be an
extremum of proper length: the calculus of variations proves that the
path is straight in the sense of zero curvature.  In terrestrial
gravity, standard $25\times 10^{-6}\unit{kg}\cdot\meter^{-1}$ cotton
sewing thread at its breaking strain of about $40\unit{N}$ would have
a sag of~$\sim 6\times 10^{-6}\meter$ over a $1\meter$ horizontal
scale: barely perceptible.  How does such a quotidian observation
translate into the vicinity of a black hole?  Would the ``warping of
space'' result in {\em visibly} curved thread?

@@

\noindent 
Solutions to equation~\ref{rdashdash} correspond to a taut string
confined to Flamm's paraboloid~\cite{flamm1916}; see
Figure~\ref{flamm}.  It is interesting to compare
equation~\ref{rdashdash} with the corresponding equation for null
geodesics, which is usually given~\cite{wald} in terms of $u=2M/r$ as
$u''=3u^2/2-u$, but is equivalent
to~$r''\left(\phi\right)=r-3M+2\left(r'\right)^2/r$.  This shows that
light, being sensitive to the time component of the metric, behaves
differently from string, which is not (at this point it might be worth
``reminding'' students that a straight line in flat space,
$r=r_0\sec\phi$, obeys $r''=r+2\left(r'\right)^2/r$).

@@

Light inextensible strings under tension adopt non-trivial
configurations including self-intersecting paths.  The path of such
strings interprets the concept ``straight line'' in general
relativity, others being null geodesics and world lines of massive
objects.  Comparing light paths with that of nearby taut strings thus
furnishes another insight into path curvature, complementing the
analysis of Abramowicz~\cite{abramowicz1992}.

In Figure~\ref{closest_approach_non_self_intersecting}, one's
classical intuition, and d'Alembert's principle, would suggest that
the black hole should accelerate to the left.  Note that the strings
are already ``straight'' in sense that they are minimal length; so
altering the tension would not change their path.  However, it is not
clear by what mechanism the required force would be transmitted: the
situation is steady, so no gravitational waves are produced, and the
string is inextensible so stores no energy.  Further work might
include analysis of taut strings near a spinning black hole.

%%Vancouver style references.

>>>>

A light string's being under tension means that its path will be an
extremum of proper length.  In flat space, the calculus of variations
proves that the path is straight in the sense of zero curvature.  In
terrestrial gravity, standard $25\times
10^{-6}\unit{kg}\cdot\meter^{-1}$ cotton sewing thread at its breaking
strain of about $40\unit{N}$ would have a sag of~$\sim 6\times
10^{-6}\meter$ over a $1\meter$ horizontal scale: barely perceptible.
How does such a quotidian observation translate into the vicinity of a
black hole?  Would the ``warping of space'' result in {\em visibly}
curved thread?

@@

\noindent 
Solutions to equation~\ref{rdashdash} correspond to a taut string
confined to Flamm's paraboloid~\cite{flamm1916}; see
Figure~\ref{flamm}.  It is interesting to compare
equation~\ref{rdashdash} with the corresponding equation for null
geodesics, which is usually given~\cite{wald} in terms of $u=2M/r$ as
$u''=3u^2/2-u$, but is equivalent
to~$r''\left(\phi\right)=r-3M+2\left(r'\right)^2/r$.  This shows that
light, being sensitive to the time component of the metric, behaves
differently from string, which is not (at this point it might be worth
``reminding'' students that a straight line in flat space,
$r=r_0\sec\left(\phi-\phi_0\right)$, obeys
$r''=r+2\left(r'\right)^2/r$).

@@

The concept of ``straight line'' is problematic in general relativity.
Although photon paths provide a definition of straightness, their
world lines are sensitive to the time component of the metric.  Light
taut strings respond only to the spatial components of the metric, and
their configuration is arguably a more intuitive and helpful way to
think about straightness.  Non-trivial configurations are adopted by
such strings, including self-intersecting paths.  Taut strings are
arguably as fundamental a definition of straighn

Comparing light paths with that of nearby taut strings thus
furnishes another insight into path curvature, complementing the
analysis of Abramowicz~\cite{abramowicz1992}.
